---
title: "RepData_PA1"
author: "zcky"
date: "6/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment Instructions
1.Code for reading in the dataset and/or processing the data  
2.Histogram of the total number of steps taken each day  
3.Mean and median number of steps taken each day  
4.Time series plot of the average number of steps taken  
5.The 5-minute interval that, on average, contains the maximum number of steps  
6.Code to describe and show a strategy for imputing missing data  
7.Histogram of the total number of steps taken each day after missing values are imputed  
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends  
9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

## Step 1  
### Code for reading in the dataset and/or processing the data  
Unzip the data and read the csv file into a variable. Explore the data.
```{r}
unzip("./activity.zip")
activityData <- read.csv("./activity.csv")
summary(activityData)
names(activityData)
head(activityData)
```

## Step 2  
### Histogram of the total number of steps taken each day  

```{r}
stepsPerDay <- aggregate(steps ~ date, activityData, sum, na.rm=TRUE)
png("plot1.png")
hist(stepsPerDay$steps)
dev.off()
hist(stepsPerDay$steps)
meanStepsPerDay <- mean(stepsPerDay$steps)
medianStepsPerDay <- median(stepsPerDay$steps)
```

## Step 3  
### Mean and median number of steps taken each day

```{r}
meanStepsPerDay <- mean(stepsPerDay$steps)
meanStepsPerDay
medianStepsPerDay <- median(stepsPerDay$steps)
medianStepsPerDay
```

## Step 4  
### Time series plot of the average number of steps taken
```{r}
stepsPerInterval<-aggregate(steps~interval, data=activityData, mean, na.rm=TRUE)
png("plot2.png")
plot(steps~interval, data=stepsPerInterval, type="l")
dev.off()
plot(steps~interval, data=stepsPerInterval, type="l")
```

## Step 5  
### The 5-minute interval that, on average, contains the maximum number of steps
```{r}
intervalWithMaxNbSteps <- stepsPerInterval[which.max(stepsPerInterval$steps),]$interval
intervalWithMaxNbSteps
```

## Step 6  
### Code to describe and show a strategy for imputing missing data  
Fill in all the missing values in the dataset with the mean per interval.  
The new data set with no missing values is contained in the variable activityDataNoNA.  
The mean didn’t change after the replacements of NAs, the median changed about 0.1% of the original value.  
```{r}
totalValuesMissings <- sum(is.na(activityData$steps))
totalValuesMissings

getMeanStepsPerInterval<-function(interval){
    stepsPerInterval[stepsPerInterval$interval==interval,]$steps
}
activityDataNoNA<-activityData

for(i in 1:nrow(activityDataNoNA)){
    if(is.na(activityDataNoNA[i,]$steps)){
        activityDataNoNA[i,]$steps <- getMeanStepsPerInterval(activityDataNoNA[i,]$interval)
    }
}
```

## Step 7  
### Histogram of the total number of steps taken each day after missing values are imputed
```{r}
totalStepsPerDayNoNA <- aggregate(steps ~ date, data=activityDataNoNA, sum)
png("plot3.png")
hist(totalStepsPerDayNoNA$steps)
dev.off()
hist(totalStepsPerDayNoNA$steps)
meanStepsPerDayNoNA <- mean(totalStepsPerDayNoNA$steps)
medianStepsPerDayNoNA <- median(totalStepsPerDayNoNA$steps)
```

## Step 8  
### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
activityDataNoNA$date <- as.Date(strptime(activityDataNoNA$date, format="%Y-%m-%d"))
activityDataNoNA$day <- weekdays(activityDataNoNA$date)
for (i in 1:nrow(activityDataNoNA)) {
    if (activityDataNoNA[i,]$day %in% c("Saturday","Sunday")) {
        activityDataNoNA[i,]$day<-"weekend"
    }
    else{
        activityDataNoNA[i,]$day<-"weekday"
    }
}
stepsByDay <- aggregate(activityDataNoNA$steps ~ activityDataNoNA$interval + activityDataNoNA$day, activityDataNoNA, mean)
names(stepsByDay) <- c("interval", "day", "steps")
library(lattice)
png("plot4.png")
xyplot(steps ~ interval | day, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
dev.off()
xyplot(steps ~ interval | day, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```

